---
title: "Lecture 10: Live code day: functions and iteration!"
author: "Wenna Xi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

## Read in the squirrels data

```{r, message = TRUE}
library(here)
library(tidyverse)

file_name <- list.files(here("Squirrel_Data"), full.names = TRUE)

# option 1
nyc_squirrels <- NULL

for (i in file_name) {
  data_temp <- read_csv(i)
  nyc_squirrels <- rbind(nyc_squirrels, data_temp)
}


#option 2
nyc_squirrels <- map(file_name, read_csv) |>
  bind_rows()

summary(nyc_squirrels)
```


## Calculate the number of squirrels seen by day of the week

```{r}
nyc_squirrels <- nyc_squirrels |>
  mutate(date = mdy(date),
         day_of_week = wday(date, label = TRUE))

nyc_squirrels |>
  select(date, day_of_week)

nyc_squirrels |>
  group_by(day_of_week) |>
  summarise(n = n())
```

## Calculate the percentage of squirrels running
```{r}
nyc_squirrels |>
  summarise(pct_running = sum(running == TRUE) / n())
```


## Plot the percentage of squirrels doing each activity
```{r}
pct_squirrels <- function(x) {
  sum(x == TRUE) / length(x)
}

nyc_squirrels |>
  summarise(across(running:foraging, pct_squirrels)) |>
  pivot_longer(cols = everything(),
               names_to = "activity", 
               values_to = "percent") |>
  mutate(activity = str_to_title(activity)) |>
  ggplot(aes(x = activity,
             y = percent * 100)) +
    geom_bar(stat = "identity") +
  labs(x = "Activity",
       y = "Percentage") +
  scale_y_continuous(labels = function(x) paste0(x, "%"))
```

## Plot the percentage of squirrels doing each activity by primary fur color
```{r}
nyc_squirrels |>
  mutate(primary_fur_color = if_else(is.na(primary_fur_color), "Missing", primary_fur_color)) |>
  group_by(primary_fur_color) |>
  summarise(across(running:foraging, pct_squirrels)) |>
  pivot_longer(cols = running:foraging,
               names_to = "activity", 
               values_to = "percent") |>
  mutate(activity = str_to_title(activity)) |>
  mutate(activity = factor(activity, levels = c("Running", "Chasing", "Climbing", "Foraging", "Eating"))) |>
  ggplot(aes(x = activity,
             y = percent * 100,
             fill = primary_fur_color)) +
    geom_bar(stat = "identity", position = "dodge", color = "gray") +
  labs(x = "Activity",
       y = "Percentage",
       fill = "Primary Fur Color") +
  scale_y_continuous(labels = function(x) paste0(x, "%"))
```
## Write a function that can plot the percentage of squirrels doing each activity by an arbitrary group (e.g., primary fur color, highlight fur color, age, date, day of week)
```{r}
varname <- "date"

plot_squirrels_by_group <- function(varname) {
  temp_data <- nyc_squirrels |>
    rename(group_var = all_of(varname))
    
    if (is.character(temp_data$group_var)) {
      temp_data <- temp_data |>
        mutate(group_var = if_else(is.na(group_var) | group_var == "?", "Missing", group_var))
    }
  
  temp_data <- temp_data|>
  group_by(group_var) |>
  summarise(across(running:foraging, pct_squirrels)) |>
  pivot_longer(cols = running:foraging,
               names_to = "activity", 
               values_to = "percent") |>
  mutate(activity = str_to_title(activity)) |>
  mutate(activity = factor(activity, levels = c("Running", "Chasing", "Climbing", "Foraging", "Eating"))) |>
    mutate(group_var = as.factor(group_var))
  
  temp_data |>
  ggplot(aes(x = activity,
             y = percent * 100,
             fill = group_var)) +
    geom_bar(stat = "identity", position = "dodge", color = "gray") +
  labs(x = "Activity",
       y = "Percentage",
       fill = str_to_title(gsub("_", " ", varname))) +
  scale_y_continuous(labels = function(x) paste0(x, "%"))
}


plot_squirrels_by_group(varname = "primary_fur_color")
plot_squirrels_by_group(varname = "highlight_fur_color")
plot_squirrels_by_group(varname = "age")
plot_squirrels_by_group(varname = "date")
plot_squirrels_by_group(varname = "day_of_week")

lapply(list("primary_fur_color", "highlight_fur_color", "age", "date", "day_of_week"), plot_squirrels_by_group)

map(list("primary_fur_color", "highlight_fur_color", "age", "date", "day_of_week"), plot_squirrels_by_group)
```


